---
title: "PCA_Clustering"
author: "RJ"
date: "8/15/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PCA and Clustering

This document summarize the learning activity using the dataset provided in [Planning Public Policy in Argentina](https://www.datacamp.com/projects/638). The original learning method is clearly explained in datacamp. A source in [kaggle](https://www.kaggle.com/gargimaheshwari/planning-public-policy-in-argentina/) might follow the original, added with the notebook owner's improvement. The kaggle source will be the learning source for this exercise.

The provided data shows the economical dan social indicators of each province. Indicators are highly correlated, which need to be confirmed by research on socioeconomical. According to writer perspective, considering the PCA explanation on [Python Machine Learning](http://diggerdnepr.ddns.net/wp-content/uploads/2019/02/python-machine-learning-2nd.pdf), feature reduction through extraction will help in clustering process by maintaining the relevant information in the original data.
>PCA helps us to identify patterns in data based on the correlationbetween features - Python Machine Learning


### Preparation
Call the required library and dataset

```{r}
library(tidyverse)
library(FactoMineR) # For PCA preprocessing
library(factoextra) # For PCA data visualization
library(ggrepel)

argentina <- read.csv(choose.files())
head(argentina)
```

The datasets has 11 variables which includes:
1. **province**: Argentina's provinces
2. **gdp**: a measure of the size of a province’s economy
3. **illiteracy**: Adult illiteracy is defined as the percentage of the population aged 15 years and over who cannot both read and write with understanding a short simple statement on his/her everyday life. According to [UNESCO](http://uis.unesco.org/en/glossary-term/adult-literacy-rate#:~:text=Adult%20illiteracy%20is%20defined%20as,on%20his%2Fher%20everyday%20life.)
4. **poverty**:  the ratio of the number of people (in a given age group) whose income falls below the poverty line
5. **deficient_infra**: 
6. **school_dropout**: rate of school drop out
7. **no_healthcare**: rate of people wihtout healthcare 
8. **birth_mortal**: birth mortality rate
9. **pop**: population
10. **movie_theatres_per_cap**: 
11. **doctors_per_cap**: ratio of doctors and population

To measure the province economic condition, GDP per capita is more relevant.
PCA feature extraction will be performed by using the factominer package. As the factominer PCA process require the input format in matrix, the current individual data point will be casted. 


```{r}

argentina_matrix <- argentina %>% 
  mutate(gpd_per_capita = gdp/pop) %>% 
  select_if(is.numeric) %>% 
  as.matrix()

head(argentina_matrix)
```


### Feature Reduction

Factominer::PCA function ease the research by automatically shows the PCA biplot. The first two principal components represent around 63% of the original data variance. The cummulative  
```{r}
argentina_pca <- PCA(argentina_matrix, scale.unit = T)
fviz_eig(argentina_pca)
```


### Clustering
At this point, clustering is performed by using 2 first principal component. 
```{r}
argentina_component <- tibble(pca_1 = argentina_pca$ind$coord[,1],
                              pca_2 = argentina_pca$ind$coord[,2]) 

# Clustering via kmeans algorithm
argentina_kmeans <- kmeans(argentina_component, centers = 4, nstart = 20, iter.max = 50)
argentina_kmeans
```

#### Visualize the clustering result
1. In the selected principal component 
```{r}
fviz_pca_ind(argentina_pca, 
             habillage = as.factor(argentina_kmeans$cluster))

```

2. In the initial original data
As can be seen in PCA biplot, in PCA-2, GDP and Population are in the same direction. Plotting will be performed by using GDP, Population and GPD per capita information. 
```{r}
argentina %>% 
  mutate(cluster = as.factor(argentina_kmeans$cluster)) %>% 
  ggplot(aes(x = cluster,
                 y = gdp/1000000,
                 color = cluster)) +
  geom_point() +
  geom_text_repel(aes(label = province), show.legend = FALSE) + #to use this, aes should be in ggplot, not in geom_point
  labs(x = "Cluster",
       y = "GDP in million",
       title = "Argentina's GDP vs Province Clusters") 
```

```{r argentina, echo=FALSE, eval=FALSE}
argentina %>% 
  mutate(cluster = as.factor(argentina_kmeans$cluster)) %>% 
  ggplot(aes(x = cluster,
                 y = gpd_per_capita,
                 color = cluster)) +
  geom_point() +
  geom_text_repel(aes(label = province), show.legend = FALSE) + #to use this, aes should be in ggplot, not in geom_point
  labs(x = "Cluster",
       y = "GDP per capita",
       title = "Argentina's GDP per capita vs Province Clusters") 
```

```{r}
argentina %>% 
  mutate(cluster = as.factor(argentina_kmeans$cluster)) %>% 
  ggplot(aes(x = cluster,
                 y = poverty,
                 color = cluster)) +
  geom_point() +
  geom_text_repel(aes(label = province), show.legend = FALSE) + #to use this, aes should be in ggplot, not in geom_point
  labs(x = "Cluster",
       y = "Poverty Rate",
       title = "Argentina's Poverty vs Province Clusters") 

```

Findings:

1. PCA-1 describes the economic condition of the province. Negative direction means the province has a good economic condition, while the positive means the province tends to have a low GPD per capita or high poverty rate


```{r}

```

